---
title: Module 11
author: "Owen Harlacker"
date: June 17, 2023
output: html_document
---


## Temperature dataset
**Read the data**
```{r}
library(data.table)
library(ggplot2)

temp <- fread('./Temperature.csv')
head(temp)

```


**Winter observations:**
```{r}
winter <- temp[Season == 'winter']
head(winter)
```


**Winter and NC:**
```{r}
winterNC <- temp[Season == 'winter' & Area == 'NC']
head(winterNC)
```

**Area, Season, and Temperature columns**
```{r}
ast <- temp[,.(Area, Season, Temperature)]
head(ast)
```

**Area, and temperature for winter**
```{r}
atWinter <- temp[Season == 'winter', .(Area, Temperature)]
head(atWinter)
```
**Observations in winter**
```{r}
obsWinter <- nrow(temp[Season=='winter'])
obsWinter

```


**Mean of temperature and salinity in winter**
```{r}
mtsw <- temp[Season == 'winter', .(m_temp = mean(Temperature, na.rm = TRUE), m_sal = mean(Salinity, na.rm=TRUE))]
mtsw
```

**Number of observations by station in the winter**
```{r}
sta_win <- temp[Season == 'winter', .N , by= Station]
sta_win
```

**Number of observations per station per season**
```{r}
obs_staSeason <- temp[,.N, by=.(Station, Season)]
obs_staSeason
```


**Aerage temperatures by month**
```{r}
m_tempMo <- temp[,.(m_temp = mean(Temperature, na.rm = TRUE)), by=.(Month)]
m_tempMo
```

**Aerage temperatures by month and area**


```{r}
m_tempMoAr <- temp[,.(m_temp = mean(Temperature, na.rm = TRUE)), by=.(Month, Area)]
m_tempMoAr
```


**Associated plot**
```{r}
ggplot(m_tempMoAr, aes(x=Month, y=m_temp, color = Area))+geom_line()+facet_wrap(~Area)+ylab('Temperature')+ggtitle('Monthly temperature for each station')+scale_x_continuous(breaks = c(2,4,6,8,10,12), limits =  c(1,12))
```
