---
title: Module 9
author: "Owen Harlacker"
date: June 14, 2023
output: html_document
---

## 1. Legolas actors:
Generate 50 random heights:
```{r}
legolas <- rnorm(50, 195, 15)
aragorn <- rnorm(50, mean=180, sd=10)
gimli <- rnorm(50, mean=132, sd=15)

sample(legolas, 5)
```


## 1a. t-test
**Legolas and Aragorn:**
```{r}
t.test(legolas, aragorn, alternative = 'two.sided')
```


Based on the first t-test, we can reject the null and say that there is a significant difference between the legolas and aragorn actors.

**Legolas and Gimli:**
```{r}
t.test(legolas, gimli, alternative = 'two.sided')
```

Based on the second t-test, we can reject the null and say that there is a significant difference between the legolas and gimli actors.

## 1b. F-test
**Gimli and Legolas:**
```{r}
var.test(gimli, legolas)
```
This test produces a very high p-value, which alludes to both datasets having a non-statistically different variance.

## 2. Iris dataset testing
```{r}
library(dplyr, quietly = TRUE)
iris.setosa <- iris %>% filter(Species == 'setosa')
iris.versicolor <- iris %>% filter(Species =='versicolor')
iris.virginica <- iris %>% filter(Species == 'virginica')

```


**Correlation test for the setosa species:**
```{r}
cor.test(iris.setosa$Sepal.Length, iris.setosa$Sepal.Width)
```

This test says that there is no statistically significant difference between the correlation of the sepal length and width, and a random distribution.


**Versicolor variant**
```{r}
cor.test(iris.versicolor$Sepal.Length, iris.versicolor$Sepal.Width)

```

This test says that there is no statistically significant difference between the correlation of the sepal length and width, and a random distribution.

**Virginica Species**
```{r}
cor.test(iris.virginica$Sepal.Length, iris.virginica$Sepal.Width)

```
This test demonstrates that there is a higher chance of correlation with this species than previous species.

## 3. Deer Dataset
Import data
```{r}
deer <- read.csv('../datasets/Deer.csv')
summary(deer)
```

**Deer caught by month**
```{r}
chisq.test(table(deer$Month))
```

The data fits the distribution with normal sampling error


**TB By Farm**
```{r}
chisq.test(table(deer$Farm, deer$Tb))
```

Data fits the distribution with normal sampling error