---
title: Mod-6
author: "Owen Harlacker"
date: June 11, 2023
output: html_document
---
## A report on Iris data from Module 3.
The goal of this report is to:

- Generate a simple report on the data
- Show all visuals created

## Data import and summary:
```{r}
# Let's get some data in here:
summary(iris)
```
# Plots


*A plot of the petal length vs the width:*
```{r}
plot(iris$Petal.Length, iris$Petal.Width,
     xlab = 'Petal Length (cm)', ylab = 'Petal Width (cm)',
     main = "Fisher's Iris Dataset", pch= 16, col = 'violet')
```
--------

*One colored by species:*

```{r}
plot(iris$Petal.Length, iris$Petal.Width,
     xlab = 'Petal Length (cm)', ylab = 'Petal Width (cm)',
     main = "Fisher's Iris Dataset", pch= 16, col = iris$Species)
legend('topleft', legend = levels(iris$Species), pch = 16, col=c(1,2,3))
```


*A subset of the data that only visualizes the Setosa species:*


```{r}


plot(subset(iris, Species == 'setosa')$Petal.Length, subset(iris, Species == 'setosa')$Petal.Width,
     xlab = 'Petal Length (cm)', ylab = 'Petal Width (cm)',
     main = "Fisher's Iris Dataset, Setosa Subset", pch= 16, col = 'goldenrod2')
```


*Differing symbology based on species:*
```{r}
plot(iris$Petal.Length, iris$Petal.Width,
     xlab = 'Petal Length (cm)', ylab = 'Petal Width (cm)',
     main = "Fisher's Iris Dataset", pch= c(15, 16, 23)[as.numeric(iris$Species)], col = 'orangered1')
legend('topleft', legend = levels(iris$Species), pch= c(15, 16, 23), col='orangered1')
```



*Symbol sizing based on sepal width:*

```{r}
plot(iris$Petal.Length, iris$Petal.Width, xlab = 'Petal Length (cm)', ylab = 'Petal Width (cm)',
     main = "Fisher's Iris Dataset", pch = 16, col = 'salmon', cex = c(0.5,1,1.5)[as.numeric(iris$Sepal.Width)])

```

*And finally one with a smoothing line*
```{r}

scatter.smooth(iris$Petal.Length, iris$Petal.Width,
               xlab = 'Petal Length (cm)', ylab = 'Petal Width (cm)',               main = "Fisher's Iris Dataset",  col = 'violet')

```
