---
title: GEOG-5680 Project 2
author: "Owen Harlacker"
date: June 21, 2023
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_floating: true
---
```{r}
library(dplyr, quietly = TRUE)
library(ggplot2, quietly=TRUE)
library(data.table, quietly=TRUE)
library(knitr)
library(kableExtra)
library(GGally)
```
## Part 1: Exploring the data

**Import the file and run some basic info functions**
```{css, echo=FALSE}
.scroll-100 {
  max-height: 100px;
  overflow-y: auto;
  background-color: inherit;
}
```
```{r}
forbes <- fread('./Forbes2000.csv')
forbes$antirank <- (nrow(forbes)-forbes$rank)

```
**A few graphs**
```{r}

ggplot(forbes, aes(x=rank, y=marketvalue)) + geom_line()+labs(x='Rank', y='Market Value', title = 'Company rank vs. market value')
```
Company types
```{r}
 ggplot(forbes, aes(x=category, fill = category))+ geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+theme(legend.position="none")+labs(x='Category', y='Number of companies', title = 'Data breakdown by category')

```


Top performer for each country
```{r}

maxCountry <- forbes[forbes[,.I[profits == max(profits, na.rm = TRUE)], by=country]$V1]
maxCountry <- maxCountry[!is.na(name)]

maxCountry %>% kable %>% kable_styling("striped", full_width = F) %>% scroll_box( height = "400px")

```

---------

```{r}

# Average profit per country
forbes %>% group_by(country) %>% summarise(meanProfit = mean(profits, na.rm = TRUE))%>% ggplot( aes(x=country,y=meanProfit, color = country))+geom_point()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5), legend.position = 'none')+labs(x='Country', y='Average profit', title = 'Average profit by country')
```


## Part 2: USA vs. Japan {.tabset}

### US Companies
```{r}
usComp <- forbes[country== 'United States']
nrow(usComp)
usComp %>% kable %>% kable_styling("striped", full_width = F) %>% scroll_box(height = "350px")

ggplot(usComp, aes(x=category, fill = category))+ geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+theme(legend.position="none")+labs(x='Category', y='Number of companies', title = 'Data breakdown by category for the US')

```
\
Banking holds the biggest category within the dataset for the US, with a not-so-close second place going to diversified financials.

### Japanese Companies
```{r}
jpComp <- forbes[country=='Japan']
nrow(jpComp)
jpComp %>% kable %>% kable_styling("striped", full_width = F) %>% scroll_box( height = "350px")

ggplot(jpComp, aes(x=category, fill = category))+ geom_bar()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+theme(legend.position="none")+labs(x='Category', y='Number of companies', title = 'Data breakdown by category for Japan', subtitle = "Note the absence of the 'Aerospace & defense' category")
```
Banking also takes the win here, with an even less close second place of Diversified Financials. Japan does not have as many clear spikes in frequency as seen with the US.


### Comparison
To find the higher ranking country, I am going to calculate the median rank for each. I will first create a boxplot of each:
```{r}
## The joys of ggplot2
ggplot(usComp, aes(x= country, y=rank, fill=country))+geom_jitter(aes(color=country))+geom_boxplot(alpha = 0.7)+geom_jitter(data=jpComp, aes(color = country))+geom_boxplot(data=jpComp, alpha =0.7 )+labs(title = 'Comparison of company rankings between the two', subtitle = 'Lower is better',x='Country', y= 'Rank')
```
\
This boxplot shows that the United States consistently ranks higher on the chart (lower rank number).










## Part 3: Creating a model of the data

Model from assets, market value, and sales
```{r}
# water reservoir %>% my sink %>% sewers %>% hopefully a treatment plant %>% the river
modelData <- forbes %>% select(c(assets, marketvalue, sales, profits))
modelData %>% ggplot(aes(x=(assets+ marketvalue+ sales), y= profits, color = 'lightsalmon'))+geom_point()
ggpairs(modelData, aes(color = 'lightsalmon'))
```
\
The correlation between the three input variables appears to be more than a bit correlated, but this could possibly be equated to the high concentration of points surrounding the origin. This will probably not have a major effect on the model.



Making the model and testing it
```{r}
profModel <- lm(profits ~ assets+ marketvalue+sales, data = modelData)
summary(profModel)
```

Looking at the summary of the model created, market value seems to have the largest effect on the calculated profit, with a ~3.6% increase on the profit per 1% increase in market value. Assets have an almost negligible negative effect on the calculation with a -0.08% estimated effect. Sales has a +0.9% effect, meaning higher sales appear to contribute to increases in profit.

This model is a pretty poor fit with an adjusted R-squared value of only 0.31. The standard error is low, which is good, but so are the t-values which isn't.

```{r}
anova(profModel)
```


Plotting the model:
```{r}

hist(residuals(profModel), col = 'slategray2', main = 'Histogram of residuals for the model', xlab = 'Residuals')
```
\
The residuals of this model are extremely biased and the distribution is skewed quite left.

#### Model Results
```{r}
# Have they made a ggplot version that can just read my mind and format the plot exactaly how I want?
modelData %>% ggplot(aes(x=(assets+ marketvalue+ sales), y= profits))+geom_point(color = '#43CD80', size = 2, )+geom_abline(slope=coef(profModel)[2], intercept = coef(profModel)[1], color='royalblue3', size = 1.5, linetype = 'dashed') + labs(x='Assets + Market Value + Sales', y='Profits', title = 'Input data and output model function')

```


Overall, this is a pretty poor model for predicting the profits. I may have spoke too quickly when I said the high correlation wasn't an issue. One of the causes of a poor model could be the variety in the way different countries handle economics, corporate regulations, etc.



## Part 4: The Model II - Making one for both the US and Japan {.tabset}

### USA

#### Input Data:
We already have the data subsetted from part 2
```{r}
usCompMod <- usComp %>% select(c(profits, assets, marketvalue, sales))
usCompMod %>% ggplot(aes(x=(assets+ marketvalue+ sales), y= profits))+geom_point(color = 'seagreen')+labs(x='Assets + Market Value + Sales', y='Profits', title = 'Input data for the US model')
ggpairs(usCompMod, aes(color = 'salmon'), title = 'Relations between all variables')
# Does ggpairs have a quiet mode?

```

#### US model creation
```{r}
usModel <- lm(profits~assets+marketvalue+sales, data = usCompMod)
summary(usModel)
anova(usModel)
hist(residuals(usModel), col = 'slategray2', main = 'Histogram of residuals for the US model', xlab = 'Residuals')
```

This model fits the data much better. The r-squared value is 0.79, the t-values are higher with much lower associated p-values. The residuals are more normally distributed, but show major bias in the ~ 1 area. This bias is countered by the second-largest group, right below 0.

#### US model plot:
```{r}
usCompMod %>% ggplot(aes(x=(assets+ marketvalue+ sales), y= profits))+geom_point(color = '#43CD80', size = 2)+geom_abline(slope=coef(usModel)[2], intercept = coef(usModel)[1], color='royalblue3', linetype = 'dashed', size = 1.5)+labs(x='Assets + Market Value + Sales', y='Profits', title = 'Input data and output model function for the US')
```


### Japan

#### Input Data:
```{r}
jpCompMod <- jpComp %>% select(c(profits, assets, marketvalue, sales))
jpCompMod %>% ggplot(aes(x=(assets+ marketvalue+ sales), y= profits))+geom_point(color = 'seagreen')
# ggpairs, the exact thing I needed but with even more confusing asthetics bindings
ggpairs(jpCompMod, aes(color = 'salmon'), title = 'Relations between all variables')
```
Japan's input data has the lowest correlation between variables of the three models. This would in theory mean a stronger, less-incorrect model.

#### Japan model creation
```{r}
japanModel <- lm(profits ~ assets + marketvalue + sales, data= jpCompMod)
summary(japanModel)
anova(japanModel)
hist(residuals(japanModel), col = 'slategray2', main = 'Histogram of residuals for the Japan Model', xlab = 'Residuals')
```
This model falls between the US and the non-subsetted model in terms of quality, with the US model being the best. The r-squared value is slightly lower at 0.68, the t-values are smaller, and the associated p-values are slightly larger. This model, however, has better residual distribution with an almost even distribution on both sides of 0.

#### Plot of the Japan Model
```{r}
# Every plot seems to require an essay length of code...
jpCompMod %>% ggplot(aes(x=(assets+ marketvalue+ sales), y= profits))+geom_point(color = '#43CD80', size = 2)+geom_abline(slope=coef(usModel)[2], intercept = coef(japanModel)[1], color='royalblue3', size = 1.5, linetype = 'dashed') + labs(x='Assets + Market Value + Sales', y='Profits', title = 'Input data and output model function for Japan')
```
